Compositional Certified Resource Bounds\ifshort{\vspace{-.15cm}}{}

### 2014-11-07 paper/main.tex
new abstract

\begin{abstract}
  The goal of quantitative resource analysis is to provide developers
  with quantitative information about the runtime behavior of software
  at development time.  Recent years have seen tremendous progress in
  automatically deriving worst-case resource bounds, yet many
  challenges in specifying, interactively deriving, formally
  certifying, and composing resource bounds remain unsolved.

  This paper describes a novel quantitative resource analysis
  framework that tackles these challenges for C programs.  The
  analysis framework consists of three parts: First, a parametric cost
  semantics formalizes the resource consumption of CompCert Clight
  programs.  Second, a quantitative Hoare logic enables users to
  interactively develop resource bounds in the Coq Proof Assistant.
  The logic is proved sound with respect to the cost semantics, and
  shallow embedding enables a derived bound to be any function that is
  definable in Coq.  Third, an automatic amortized resource analysis
  for Clight computes derivations in the quantitative Hoare logic.
  It is the first automatic amortized analysis that can derive bounds
  that depend on negative integers or differences between numbers,
  which is crucial to handle typical systems code.  Both, the
  quantitative logic and the automatic amortized analysis are
  naturally compositional and can be combined to semi-automatically
  derive global resource bounds.
  %
  An experimental evaluation demonstrates the practicality of the
  analysis framework.  The expressivity of the logic is shown by
  manually deriving customized bounds that are tailored to specific
  algorithms.  A comparison of the automatic amortized analysis with
  other automatic tools on 30 challenging loop and recursion patterns
  from the literature and open-source software shows that the bounds
  derived by the automatic amortized analysis are often more precise.
\end{abstract}

>>>>

\begin{abstract}
This paper presents a new approach to quantitative
analysis for C programs.  The technique described uses
ideas from amortized analysis and abstract interpretation
to derive tight linear worst-case bounds that come with a
proof certificate.  Contrary to most available tools for
C, ours has a compositional design that provides the
programmer with bounds expressed in terms of meaningful
variables.  The compositionality allows to handle uniformly
nested sequenced loops and recursive sequenced function
calls.  The automatic analysis is implemented
as a tool that leverages state-of-the-art SMT solvers
LP solvers to automatically infer resource bounds on
a wide range of C programs.  It is the first automatic
amortized analysis that can derive bounds that depend
on negative integers and differences between numbers.
To prove its soundness and give a concrete shape to proof
certificates, we developed a quantitative Hoare logic
that we proved sound using the Coq proof assistant.  This
logic allows manual proofs of arbitrarily complex resource
bounds on C programs and interaction with the automation.
To prove the scalability and precision of this new tool,
we run it on multiple code snippets from the literature,
reproduce results from previous work, and analyze a
consequent part of a modern benchmark suite.

%   The goal of quantitative resource analysis is to provide developers
%   with quantitative information about the runtime behavior of software
%   at development time.  Recent years have seen tremendous progress in
%   automatically deriving worst-case resource bounds, yet many
%   challenges in specifying, interactively deriving, formally
%   certifying, and composing resource bounds remain unsolved.
%
%   This paper describes a novel quantitative resource analysis
%   framework that tackles these challenges for C programs.  The
%   analysis framework consists of three parts: First, a parametric cost
%   semantics formalizes the resource consumption of CompCert Clight
%   programs.  Second, a quantitative Hoare logic enables users to
%   interactively develop resource bounds in the Coq Proof Assistant.
%   The logic is proved sound with respect to the cost semantics, and
%   shallow embedding enables a derived bound to be any function that is
%   definable in Coq.  Third, an automatic amortized resource analysis
%   for Clight computes derivations in the quantitative Hoare logic.
%   It is the first automatic amortized analysis that can derive bounds
%   that depend on negative integers or differences between numbers,
%   which is crucial to handle typical systems code.  Both, the
%   quantitative logic and the automatic amortized analysis are
%   naturally compositional and can be combined to semi-automatically
%   derive global resource bounds.
%   %
%   An experimental evaluation demonstrates the practicality of the
%   analysis framework.  The expressivity of the logic is shown by
%   manually deriving customized bounds that are tailored to specific
%   algorithms.  A comparison of the automatic amortized analysis with
%   other automatic tools on 30 challenging loop and recursion patterns
%   from the literature and open-source software shows that the bounds
%   derived by the automatic amortized analysis are often more precise.
\end{abstract}
